---
title: "Simulation Exercise (Exponential distribution)"
author: "José Antonio García Ramirez"
date: "December 31, 2016"
output: 
  md_document:
  variant: markdown_github
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, warning = FALSE )
library(ggplot2)
library(dplyr)
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
 # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                    ncol = cols, nrow = ceiling(numPlots/cols))
  }
 if (numPlots==1) {
    print(plots[[1]])
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```

## Introduction and overview: Exponential distribution
In this exercise we will perform some simulation of variables with exponential distribution and verify the CLT.

The density function of exponential distribution is given by:

$$ f(x) = \lambda e^{-\lambda x}, \forall \lambda \ge 0$$

And it's cumulative distribution function is:

$$ \int_{\lambda \ge 0} \lambda e^{-\lambda x} = 1 -e\lambda x $$

And this distribution has many applications in the survival analysis, such as lifetimes and even in industrial statistics (including software development) [1](https://en.wikipedia.org/wiki/Survival_analysis)


## Sample mean and variance

Let's first simulate a thousand random variables $x \sim Exp(\lambda = 0.2)$, let's to fix the seed for reproducibility. 


```{r }
set.seed(176325)
simula <- rexp(1000,0.2)
simula <- as.data.frame(simula)
sampleMean <- mean(simula$simula)
sampleVar <- var(simula$simula)
g1 <- ggplot(simula, aes(simula)) + geom_histogram(aes(y = ..count../sum(..count..)),
                                                      fill = 'deepskyblue2') +
    ggtitle("Simulation Exp(lambda=0.2)") + ylab('')+ xlab('') +theme_bw() + 
stat_function(fun = dexp, colour = "springgreen", args = 0.2) + 
	geom_vline( xintercept = sampleMean, color = 'violetred') +
	geom_vline( xintercept = 1/0.2, color = "red")  
g2 <- ggplot(simula, aes(simula)) + geom_histogram(aes(y = ..count../sum(..count..)),
                                                      fill = 'deepskyblue2') +
    ggtitle("Simulation Exp(lambda=0.2)") + ylab('')+ xlab('') +theme_bw() + 
stat_function(fun = dexp, colour = "springgreen", args = 0.2) + 
	geom_vline( xintercept = sampleMean, color = 'violetred') +
	geom_vline( xintercept = 1/0.2, color = "red") + xlim(c(3.5,7.5)) 
multiplot(g1,g2, cols = 2)

```

Figure 1: (Right) Distribution of the simulated sample. (Left) Zoom on the sample mean, in color violet and theoretical mean ( $\frac{1}{\lambda}=5$ ) in color red.


As we can see in figure 1, the sample mean of the thousand simulations is very close to the theoretical average `r sampleMean` as well as the sample variance `r sampleVar` Which, for this distribution, coincide and are $\frac{1}{\lambda}=5$ and $(\frac{1}{\lambda})^{2}=25$ respectively.

## Central limit theorem

Let's check the CLT by simulating a thousand averages of 40 random variables $x \sim Exp(0.2)$

```{r }
set.seed(176325)
simula <- function(lambda)
{
	sim <- rexp(40,0.2)
	return(c(mean(sim), var(sim)))
}
simulation <- mapply(simula,rep(0.2,1000))
simulation <- as.data.frame(t(simulation))
names(simulation) <- c("means", "var")
means <- mean(simulation$means)
vars <- var(simulation$means)
lambda <- .2


#simulation <- simulation %>% mutate( score = (means-beta) / ( beta /sqrt(40)))
ggplot(simulation, aes(means)) + geom_density(aes(colour = 'Simulation') )+
    ggtitle("Simulation average 40 Exp(lambda=0.2)") + ylab('')+ xlab('')  + 
stat_function(fun = dnorm, colour = "springgreen", args =  c(1/lambda, 1/lambda/sqrt(40))) +
theme_bw()+theme(legend.title=element_blank())
```

Figure 2: Distribution of a thousand averages $Exp (0.2)$

The average of the simulation is of `r means` while its standard deviation is of `r vars` so we can see in the figure 2 the simulated sample approaches well to a distribution $N (\frac{1}{\lambda}=5, \frac{ \frac{1} {\lambda}}{\sqrt {40}}=0.7905694$ ( figure 2 green).

