---
title: "Easier reactive plots "
author: "José Antonio García Ramirez"
date: "January 30, 2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache =TRUE)
```

## Previous work and context

A few weeks ago I made the following [visualization](https://foufoo.shinyapps.io/plotly/) using the packages _shiny_ and _ggvis_ at the time they were the only tools (in R) that I knew to present a reactive graph  which it to put the mouse over the data points displays aditional data (in this case an average value and a maximum value at each point).

```{r plotly, echo = FALSE, include = FALSE, warnings = FALSE}
library(plotly)
library(dplyr)
load('/home/fou/Desktop/Online/RCoursera/R-Coursera/DevDataProducts/plotly/stats.Rdata')
p <- plot_ly(
  stats2, x = ~dinero, y = ~n,
  color = ~factor(periodo), size = ~media) %>%
	layout(title = "Cost vs number of beneficiaries",
           xaxis = list(title = "Total amount"), 
           yaxis = list(title = "Max. Number of procedures"))

```

## 
The same plot is repeated using _plotly_ with the same data. 

```{r, warning= FALSE, message=  FALSE}
p
```




## Conclutions and further work

Today for the graph made with _ggvis_ I have not found a way to share the image so that the functionality of displaying information by putting the mouse over the points is maintained, I have only managed to share the image as .png, however when using _plotly_ the development time was much, much, less a web service is not required and I was able to share the graph keeping the said functionality in a .html file :D
```

