---
title: "Reproducible Research: Peer Assessment 1 (Fancy clothes)"
author: "José Antonio García Ramirez"
date: "December 11, 2016"
output:
  md_document:
    variant: markdown_github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the (smart) data

First we get the data directly from the [url](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip), then unzip it and with the library -lubridate- and transform the variable 'date' to a lubridate's date type.



```{r getData}
temp <- tempfile()
download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip',temp)
data <- read.csv(unz(temp, "activity.csv"), header = TRUE)
head(data)
unlink(temp)
library(lubridate)
data$date <- ymd(data$date)
rows <- dim(data)[1]
vars <- dim(data)[2]
```

And we check for the dimensionality of the dataset to check that it has `r rows` observations and `r vars` variables.

## Mean total number of steps taken per day.
First we observe the distribution of the variable: number of step of the anonymous individual in each time, first group by the date (I will use the package dplyr to manipulate the data).

```{r primero} 
library(ggplot2)
library(dplyr)
a <- data %>% group_by(date) %>% summarise(steps=sum(steps, na.rm =TRUE)) 
m <- mean(a$steps, na.rm = TRUE)
med <- median(a$steps, na.rm = TRUE)
rugs <- data.frame(steps=c(m,med))
ggplot(a, aes(x = steps)) + geom_histogram(aes(y = ..count../sum(..count..)), fill='green3') + ggtitle("Daily steps' distribution") + ylab('')+ xlab('') +theme_bw() +geom_rug(data = rugs, colour = c('red', 'blue')) 
```

The plot has two rugs, in red color the mean number of steps (`r m`) and another one in blue color indicating the median (`r med`).



## Average daily activity pattern.

First we observe the distribution of the variable across the intervals and it is calculated it's average number of steps

```{r segundo }
a <- data %>% group_by(interval) %>% summarise(steps=mean(steps, na.rm = TRUE)) 
maxInterval <-a$interval[which.max(a$steps)]
ggplot(a, aes(x = interval, y = steps)) + geom_line(colour = 'deepskyblue2') +
	ylab("Steps' number") + theme_bw()  + ggtitle('Average steps')+
	theme_bw()
```

The interval with maximum average steps' number is (`r maxInterval`).


## Imputing missing values.
 
```{r tercero }
nas <- sum(is.na(data$steps))
```

The number of missing values in the dataset is `r nas`. I will fill NA's with the mean of the interval. 


```{r tercero2}
means <- data  %>% group_by(interval) %>% summarise( med = mean(steps, na.rm = TRUE)) # the means of all intervals 
index <- which( is.na(data$steps))  #the localitations of NA's in the dataset
data2 <- data
data2$steps[index] <- mapply(function(x){return( means$med[ which(means$interval == x ) ])   }, data$interval[index]) #replace NA's

a <- data2 %>% group_by(date) %>% summarise(steps=sum(steps, na.rm =TRUE)) 
m <- mean(a$steps, na.rm = TRUE)
med <- median(a$steps, na.rm = TRUE)
rugs <- data.frame(steps=c(m,med))
ggplot(a, aes(x = steps)) + geom_histogram(aes(y = ..count../sum(..count..)), fill='green3') + ggtitle("Daily steps' distribution") + ylab('')+ xlab('') +theme_bw() +geom_rug(data = rugs, colour = c('red', 'blue')) 
```

The plot has two rugs, in red color the mean number of steps (`r m`) and another one in blue color indicating the median (`r med`). The effect after replace NA's is that the distribution is more simetric respect to the median also the mean and median are aproximaly equal.


## Differences in activity patterns between weekdays and weekends

```{r cuarto}
data2$end <- weekdays(data2$date) %in% c('sábado','domingo') 
data2$end <- factor(data2$end, labels = c("weekday", "weekend" )) 
a <- data2 %>% group_by(interval,end) %>% summarise(steps=mean(steps, na.rm = TRUE)) 
ggplot(a, aes(x = interval, y = steps)) + geom_line(colour = 'deepskyblue2') +
	ylab("Steps' number") + theme_bw()  + ggtitle('Average steps acroos the week')+
	theme_bw() + facet_grid(end ~ .)
```


There are differences between activity in the week and the weekend, in the week the activity has more peaks but in average the activity in weekend is bigger, also respecto to the time in the week the activity ends first that in weekend.

