---
title: "EcoBici (Shared bicycles in Mexico)"
author: "José Antonio García Ramirez"
date: "February 5, 2017"
output:
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache =TRUE)
```

## Previous work and context
In Mexico City there is a bicycle loan system, [EcoBici](https://www.ecobici.cdmx.gob.mx/), and A few months ago I do a study about them about which stations are the most popular and which routes (travel from station to station) are more used, you can consult the study [here](https://github.com/fou-foo/opi_test/blob/master/exa_opi.html) (But only available in Spanish but add a nice map made with _leaflet_). 

##  Updates to the previous work
When the study was carried out, a graph was needed to explore the flows between the stations.
The _shiny_ app is documented as follows:

   * The second tab shows a display of stations of a bicycle loan service in Mexico City, each node represents a station in the city.
   * The edges represent very popular travel routes, on the left side you can choose the threshold from which routes are drawn in which $ n > 50$ travels arrive and you can also draw the routes in which they leave more than $ m > 50 $ trips per service station.
   * The service API also provides information about the inflow of the stations, that information is what determines in node size.
   * Finally an option to draw different colors of the nodes is given at the user's choice.
   * Also on the graph you can zoom to see it better.

## Example output App

```{r grafo}
library(shiny)
library(networkD3)
setwd("/home/fou/Desktop/Online/RCoursera/R-Coursera/DevDataProducts/ShinyGraph")
load(file= "oct.Rdata")
library(dplyr)
flujo_neg <- oct %>% select(Ciclo_Estacion_Retiro, Ciclo_Estacion_Arribo) %>%
		arrange(Ciclo_Estacion_Retiro, Ciclo_Estacion_Arribo) %>%
		group_by(Ciclo_Estacion_Retiro) %>% summarise(w = n()) #viajes que salen de estacion
flujo_neg$tipo <- "Salida" 
names(flujo_neg) <- c('Estacion',"wins", "tipo")
flujo_pos <- oct %>% select(Ciclo_Estacion_Retiro, Ciclo_Estacion_Arribo) %>%
		arrange( Ciclo_Estacion_Arribo, Ciclo_Estacion_Retiro) %>%
		group_by(Ciclo_Estacion_Arribo) %>% summarise(dentro = n()) %>% arrange(desc(dentro))
flujo_pos$tipo <- "Entrada"
names(flujo_pos) <- c('Estacion',"w","tipo")
edges_in <- oct %>% select(Ciclo_Estacion_Retiro, Ciclo_Estacion_Arribo) %>%
			group_by(Ciclo_Estacion_Retiro, Ciclo_Estacion_Arribo) %>% summarise(w = n()) %>%
			arrange(desc(w))
ins <- subset(edges_in, w > 200)
edges_out <- oct %>% select( Ciclo_Estacion_Arribo, Ciclo_Estacion_Retiro) %>%
	group_by( Ciclo_Estacion_Arribo, Ciclo_Estacion_Retiro) %>% summarise(w = n()) %>%
	arrange(desc(w))
out <- subset(edges_out, w > 200)
aristas <- rbind(ins, out)
aristas <- as.data.frame(aristas)
nodos <- as.data.frame(flujo_pos)
names(aristas) <- c("source", "target", "value" )
names(nodos) <- c("id", "w", "tipo")
aristas$source <- aristas$source - 1
aristas$target <- aristas$target - 1
nodos$id <- nodos$id -1 
aristas$value <- scale(aristas$value, scale = FALSE)
forceNetwork(Links = aristas, Nodes = nodos,
					 Source = 'source', Target = 'target',
					 Value = "value", opacity = 0.8,
					 NodeID = 'id', Group = "w",
					 radiusCalculation = "Math.pow(Math.log(d.nodesize), 1.5)",
					 Nodesize  = "w", 
					 zoom = TRUE
					 #legend = TRUE
		)
```




## Conclutions and further work

[app](https://foufoo.shinyapps.io/appgrafo/)

  * Search in the _graph_
  * Custom _graph_ maybe with _igraph_ package
  * Deploy own visualization in R
  * Change language ...